Cluster Difference Imaging Photometric Survey. III. Rotation and Lithium in an Open Cluster Spanning 500\,Parsecs

### 2021-04-08 cluster_paper/ms.tex
update ms

\title{
  Cluster Difference Imaging Photometric Survey. III.
  An Open Cluster Spanning 500\,Parsecs
}

@@

Star clusters form when dense filaments in hierarchically structured
molecular clouds collide and collapse \citep{shu_star_1987}.  Over the
first 10\,Myr, feedback effects including protostellar outflows,
photoionization, radiation pressure, and supernova shocks disperse the
gas out of the cloud \citep{krumholz_star_2019}.  Since only $\sim
1\%$ of the cloud mass is converted into stars, gas dispersal enables
the majority ($\sim$90\%) of stars in the cluster to escape the
cluster's gravitational well \citep{lada_embedded_2003}.

@@

The detection of dispersed stellar associations, although possible
before Gaia, has now reached a breakthrough pace \citep[{\it
e.g.},][]{de_zeeuw_hipparcos_1999,bergond_gravitational_2001,zuckerman_young_2004,oh_comoving_2017,cantatgaudin_gaia_2018,gagne_banyan_XII_2018,gagne_banyan_XIII_2018,kounkel_apogee2_2018,zari_3d_2018,kounkel_untangling_2019,furnkranz_comabernbhr_2019}.
The populations found by these studies can be summarized as follows.
On one end are coeval groups with no discernable cores ({\it e.g.},
Psc Eri and $\mu$ Tau,
\citealt{meingast_psceri_2019,curtis_tess_2019,gagne_mutau_2020}).  On
the other are hierarchically structured associations with many over
and under-densities  \citep[{\it e.g.} the Sco-Cen and Vela
associations][]{pecaut_star_2016,cantatgaudin_velaOB2_2019}.  Here, we
focus on an intermediate regime: low-density halos associated with a
single overdensity, typically an open cluster that was known before
Gaia
\citep[see][]{kounkel_untangling_2019,kounkel_untanglingII_2020,meingast_2021}.
In some cases, these halos could correspond to tidal tails, as have
been reported in the Hyades
\citep{meingast_hyades_2019,roser_hyades_2019}, the Ursa Major moving
group \citep{gagne_lowmassUMA_2020}, and in Coma Berenices
\citep{tang_comaber_2019}.

@@

We have recently been making TESS light curves of stars reported to be
members of coeval populations (CDIPS, \citealt{bouma_cdipsI_2019}).
Our analysis of TESS Sectors 1-13 yielded light curves of
\numsouthernuniqlcs\ candidate cluster members in the Southern
Ecliptic hemisphere, which are available on
MAST\footnote{\url{https://archive.stsci.edu/hlsp/cdips}}.

@@

Section~\ref{sec:gaia} summarizes the astrometric analyses of the Gaia
data that led to our interest in \cn.  Section~\ref{sec:agedate}
measures the age of the cluster's halo and core using Gaia photometry
(Section~\ref{subsec:hr}), TESS gyrochronology
(Section~\ref{subsec:tess}), and lithium depletion
(Section~\ref{subsec:lithium}).  In Section~\ref{sec:discussion} we
discuss the implications of our analysis for NGC\,2516 specifically
and stellar spin-down and open cluster evolution generally.
Section~\ref{sec:conclusion} gives our conclusions.

@@

\subsection{Rotation at 100-200\,Myr}
\label{disc:rotn}

@@

Comparing the explanations of disk-locking against tidal
synchronization, the population statistics seem to strain the
possibility of tidal synchronization.  A quarter of the \cn\ members
are rapidly rotating.  In comparison, in the field half of Sun-like
stars are binaries, and $\approx$9\% have periods below 100\,days
\citep{raghavan_survey_2010}.  If we assume that all binaries with
sub-100\,day periods become tidally synchronized, we can only explain
a rapid rotator occurrence rate of $\approx$5\%.  Elevated binarity
fractions in pre-MS stars likely do not change this picture \citep[see
Sec.~4.4 of][]{duchene_stellar_2013}.

>>>>

\title{
  Cluster Difference Imaging Photometric Survey. III.
  Rotation and Lithium in an Open Cluster Spanning 500\,Parsecs
}

@@

Star clusters form when dense filaments in hierarchically structured
molecular clouds collide and collapse \citep{shu_star_1987}.  Over the
first 10\,Myr, feedback effects including protostellar outflows,
photoionization, radiation pressure, and supernova shocks disperse the
gas out of the cloud \citep{krumholz_star_2019}.  Since only
$\sim$$1\%$ of the cloud mass is converted into stars, gas dispersal
enables the majority ($\sim$90\%) of stars in the cluster to escape
the cluster's gravitational well \citep{lada_embedded_2003}.

@@

The detection of dispersed stellar associations, although possible
before Gaia, has now reached a breakthrough pace \citep[{\it
e.g.},][]{de_zeeuw_hipparcos_1999,bergond_gravitational_2001,zuckerman_young_2004,oh_comoving_2017,cantatgaudin_gaia_2018,gagne_banyan_XII_2018,gagne_banyan_XIII_2018,kounkel_apogee2_2018,zari_3d_2018,kounkel_untangling_2019,furnkranz_comabernbhr_2019}.
The populations found by these studies can be summarized as follows.
On one end are coeval groups with no discernable cores ({\it e.g.},
Psc Eri and $\mu$ Tau,
\citealt{meingast_psceri_2019,curtis_tess_2019,gagne_mutau_2020}).  On
the other are hierarchically structured associations with many over
and under-densities  \citep[{\it e.g.} the Sco-Cen and Vela
associations,][]{pecaut_star_2016,cantatgaudin_velaOB2_2019}.  Here, we
focus on an intermediate regime: low-density halos associated with a
single overdensity, typically an open cluster that was known before
Gaia
\citep[see][]{kounkel_untangling_2019,kounkel_untanglingII_2020,meingast_2021}.
In some cases, these halos could correspond to tidal tails, as have
been reported in the Hyades
\citep{meingast_hyades_2019,roser_hyades_2019}, the Ursa Major moving
group \citep{gagne_lowmassUMA_2020}, and in Coma Berenices
\citep{tang_comaber_2019}.

@@

We have recently been making TESS light curves of stars reported to be
members of coeval populations (CDIPS, \citealt{bouma_cdipsI_2019}) and
searching them for planets \citep{bouma_cluster_2020}.
Our analysis of TESS Sectors 1-13 yielded light curves of
\numsouthernuniqlcs\ candidate cluster members in the Southern
Ecliptic hemisphere, which are available on
MAST\footnote{\url{https://archive.stsci.edu/hlsp/cdips}}.

@@

Section~\ref{sec:gaia} summarizes the astrometric analyses of the Gaia
data that led to our interest in \cn.  Section~\ref{sec:agedate}
measures the age of the cluster's halo and core using Gaia photometry
(Section~\ref{subsec:hr}), TESS gyrochronology
(Section~\ref{subsec:tess}), and relative lithium abundances
(Section~\ref{subsec:lithium}).  In Section~\ref{sec:discussion} we
discuss the implications of our analysis for NGC\,2516 specifically
and stellar spin-down more generally.
Section~\ref{sec:conclusion} gives our conclusions.

@@

\subsection{Stellar Rotation at 100-200\,Myr}
\label{disc:rotn}

@@

Comparing the explanations of disk-locking against tidal
synchronization, the population statistics seem to strain the
possibility of tidal synchronization.  A quarter of the \cn\ members
are rapidly rotating.  In comparison, in the field half of Sun-like
stars are binaries, and $\approx$9\% of these binaries have periods below 100\,days
\citep{raghavan_survey_2010}.  If we assume that all binaries with
sub-100\,day periods become tidally synchronized, then we can only explain
a rapid rotator occurrence rate of $\approx$5\%.  Elevated binarity
fractions in pre-MS stars likely do not change this picture \citep[see
Sec.~4.4 of][]{duchene_stellar_2013}.
